---
output: rmarkdown::github_document
---

# Potential bug in the R package `Synth`

Running the first example in the docs of `Synth::synth()` (using the “Toy panel dataset”) gives:

```{r}
library(Synth)
packageVersion("Synth")

# load data
data(synth.data)

# create matrices from panel data that provide inputs for synth()
dataprep.out <-
 dataprep(
   foo = synth.data,
   predictors = c("X1", "X2", "X3"),
   predictors.op = "mean",
   dependent = "Y",
   unit.variable = "unit.num",
   time.variable = "year",
   special.predictors = list(
     list("Y", 1991, "mean"),
     list("Y", 1985, "mean"),
     list("Y", 1980, "mean")
   ),
   treatment.identifier = 7,
   controls.identifier = c(29, 2, 13, 17, 32, 38),
   time.predictors.prior = c(1984:1989),
   time.optimize.ssr = c(1984:1990),
   unit.names.variable = "name",
   time.plot = 1984:1996
 )

synth.out <- synth(dataprep.out)

gaps_orig <- dataprep.out$Y1plot - (
  dataprep.out$Y0plot %*% synth.out$solution.w
)
```
 


Now, I change the order of the predictors to be `c("X3", "X1", "X2")` without changing anything else:

```{r}
dataprep.out <-
 dataprep(
   foo = synth.data,
   predictors = c("X3", "X1", "X2"),
   predictors.op = "mean",
   dependent = "Y",
   unit.variable = "unit.num",
   time.variable = "year",
   special.predictors = list(
     list("Y", 1991, "mean"),
     list("Y", 1985, "mean"),
     list("Y", 1980, "mean")
   ),
   treatment.identifier = 7,
   controls.identifier = c(29, 2, 13, 17, 32, 38),
   time.predictors.prior = c(1984:1989),
   time.optimize.ssr = c(1984:1990),
   unit.names.variable = "name",
   time.plot = 1984:1996
 )

synth.out <- synth(dataprep.out)

gaps_mod <- dataprep.out$Y1plot - (
  dataprep.out$Y0plot %*% synth.out$solution.w
)
```

Comparing the two:

```{r}
out <- cbind(gaps_orig, gaps_mod)
out

matplot(out, type = "b", pch = 1, col = 1:2)
legend("topleft", legend = c("orig", "mod"), col = 1:2, pch = 1)
```

Not a big difference in this case, but still a difference whereas there shouldn't be any. It isn't mentioned anywhere in the documentation that the order of controls is important (and I don't know why it should be like that).